---
title: "R Notebook"
output: html_notebook
---


```{r}
cmmice = showmetrics(MICEpreds, outcome)

cmmice = confusionMatrix(as.factor( round(MICEpreds)),as.factor(outcome), positive='1')
cmmice
cmmice$overall[1]
```


```{r}
cmmice = confusionMatrix(as.factor( round(MICEpreds)),as.factor(outcome), positive='1')
cmmr = confusionMatrix(as.factor( round(MRpreds)),as.factor(outcome), positive='1')
cmmidas = confusionMatrix(as.factor( round(MIDASpreds)),as.factor(outcome), positive='1')
cmmidasvae = confusionMatrix(as.factor( round(MIDASVAEpreds)),as.factor(outcome), positive='1')
cmcc = confusionMatrix(as.factor( round(CCpreds)),as.factor(CCoutcome), positive='1')
# cmmice$byClass["Pos Pred Value"]
#   overall["Accuracy"]


method = c("Complete case", "MICE", "missRanger", "MIDAS", "MIDASVAE")
metric = c("Accuracy", "Brier", "Pos Pred Value", "Neg Pred Value", "Mean Loss", "AUC")
accuracy = c(cmcc$overall["Accuracy"],cmmice$overall["Accuracy"],cmmr$overall["Accuracy"],cmmidas$overall["Accuracy"]
             ,cmmidasvae$overall["Accuracy"])
accuracy = accuracy/accuracy[1] * 100

Brier = c(CCbrier, MICEbrier, MRbrier, MIDASbrier, MIDASVAEbrier)
Brier = Brier/Brier[1] * 100

posPredValue = c(cmcc$byClass["Pos Pred Value"],cmmice$byClass["Pos Pred Value"],cmmr$byClass["Pos Pred Value"]
                 ,cmmidas$byClass["Pos Pred Value"],cmmidasvae$byClass["Pos Pred Value"])
posPredValue =   posPredValue/posPredValue[1] * 100

negPredValue = c(cmcc$byClass["Neg Pred Value"],cmmice$byClass["Neg Pred Value"],cmmr$byClass["Neg Pred Value"]
                 ,cmmidas$byClass["Neg Pred Value"],cmmidasvae$byClass["Neg Pred Value"])
negPredValue = negPredValue/negPredValue[1] * 100

meanloss = c(CCmeanloss, MICEmeanloss, MRmeanloss, MIDASmeanloss, MIDASVAEmeanloss)
meanloss = meanloss/meanloss[1] * 100

auc = c(CCci[2], MICEci[2],MRci[2], MIDASci[2], MIDASVAEci[2])
auc = auc/auc[1] * 100
# CCci[1]
# CCci[2]
# CCci[3]
# CCci
#summarydf = data.frame(method, accuracy, accuracyN, precision, precisionN, negpredvalue, negpredvalueN, meanloss, meanlossN, auc,
summarydf = data.frame(method, accuracy, Brier, posPredValue, negPredValue, meanloss, auc)

```


```{r}
library(reshape)
library(data.table)

sdf2 = data.frame(t(summarydf))
sdf2 

names(sdf2) = summarydf$method
setDT(sdf2, keep.rownames = TRUE)[]
colnames(sdf2)[1] <- "Metric"

#sdf3 = sdf2[c(3,5,7,9, 11),]
sdf3 = sdf2[-1]
sdf3$"Complete case"=round(as.numeric(as.character(sdf3$"Complete case")),1)
sdf3$"MICE"=round(as.numeric(as.character(sdf3$"MICE")),1)
sdf3$"missRanger"=round(as.numeric(as.character(sdf3$"missRanger")),1)
sdf3$"MIDAS"=round(as.numeric(as.character(sdf3$"MIDAS")),1)
sdf3$"MIDASVAE"=round(as.numeric(as.character(sdf3$"MIDASVAE")),1)
sdf3
```

```{r}
library(reshape2)

# reshape your data into long format
sdf3l <- melt(sdf3, id=c("Metric"))
sdf3l

#create color palette:
library(RColorBrewer)
coul = brewer.pal(5, "Dark2") 

#values = c("red","blue","green","yellow"), 
# make the plot
ggplot(sdf3l) +
  geom_bar(aes(x = Metric, y = value, fill = variable), 
           stat="identity", position = "dodge", width = 0.7) +
  scale_fill_manual("Method\n", values = coul, 
                    labels = c("Complete Case", "MICE", "missRanger", "MIDAS", "MIDASVAE")) +
  labs(x="",y="Relative Performance %\n") +
  ggtitle("Performance Relative to Complete Case (100%)") +
  ylim(c(0,160)) +
  theme_bw(base_size = 14) + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r}
# CCpredsMICE <- predict.glm(fitCC, newdata = dfMICEtestPrepped[dfMICEtestPrepped$.imp==1,], type="response")
# CCpredsMR <- predict.glm(fitCC, newdata = dfMRtestPrepped[dfMRtestPrepped$.imp==1,], type="response")
# CCpredsMIDAS <- predict.glm(fitCC, newdata = dfMIDAStestPrepped[dfMIDAStestPrepped$.imp==1,], type="response")
# CCpredsMIDASVAE <- predict.glm(fitCC, newdata = dfMIDASVAEtestPrepped[dfMIDASVAEtestPrepped$.imp==1,], type="response")

#applying imputed test datasets to Complete Case model i.e. more data
CCMICEpreds <- rowMeans(sapply(1:m, function(x) predict.glm(fitCC, newdata = dfMICEtestPrepped[dfMICEtestPrepped$.imp==x,], type="response")))
CCMRpreds <- rowMeans(sapply(1:m, function(x) predict.glm(fitCC, newdata = dfMRtestPrepped[dfMRtestPrepped$.imp==x,], type="response")))
CCMIDASpreds <- rowMeans(sapply(1:m, function(x) predict.glm(fitCC, newdata = dfMIDAStestPrepped[dfMIDAStestPrepped$.imp==x,], type="response")))
CCMIDASVAEpreds <- rowMeans(sapply(1:m, function(x) predict.glm(fitCC, newdata = dfMIDASVAEtestPrepped[dfMIDASVAEtestPrepped$.imp==x,], type="response")))


CCMICElossvector = ifelse(outcome==1, yes=-log(CCMICEpreds), no=-log(1-CCMICEpreds))
CCMRlossvector = ifelse(outcome==1, yes=-log(CCMRpreds), no=-log(1-CCMRpreds))
CCMIDASlossvector = ifelse(outcome==1, yes=-log(CCMIDASpreds), no=-log(1-CCMIDASpreds))
CCMIDASVAElossvector = ifelse(outcome==1, yes=-log(CCMIDASVAEpreds), no=-log(1-CCMIDASVAEpreds))

CCMICEmeanloss = mean(CCMICElossvector) 
CCMICEsumloss = sum(CCMICElossvector)
CCMRmeanloss = mean(CCMRlossvector) 
CCMRsumloss = sum(CCMRlossvector)
CCMIDASmeanloss = mean(CCMIDASlossvector) 
CCMIDASsumloss = sum(CCMIDASlossvector)
CCMIDASVAEmeanloss = mean(CCMIDASVAElossvector) 
CCMIDASVAEsumloss = sum(CCMIDASVAElossvector)

cat("mean loss\n")
cbind(MICEloss = CCMICEmeanloss, MRloss = CCMRmeanloss, MIDASloss = CCMIDASmeanloss, MIDASVAEloss = CCMIDASVAEmeanloss, CCLoss=CCmeanloss)

cat("\nsum loss\n")
cbind(MICEloss = CCMICEsumloss, MRloss = CCMRsumloss, MIDASloss = CCMIDASsumloss, MIDASVAEloss = CCMIDASVAEsumloss, CCLoss=CCsumloss)

```
The above uses averaged probability over all 40 datasets
Compared to using just 1st imputed dataset
mean loss
      MICEloss    MRloss CCMIDASmeanloss MIDASVAEloss    CCLoss
[1,] 0.3246275 0.3673971       ?    0.3832039 0.1696779

sum loss
     MICEloss   MRloss MIDASloss MIDASVAEloss  CCLoss
[1,] 3836.123 4341.531  4693.304     4528.321 855.007


Compared to using imputed models
mean loss
      MICEloss    MRloss MIDASloss MIDASVAEloss    CCLoss adjustedCCLoss
[1,] 0.2297861 0.2604949 0.2397141    0.2334464 0.1696779      0.4167898

sum loss
     MICEloss   MRloss MIDASloss MIDASVAEloss  CCLoss adjustedCCLoss
[1,] 2715.382 3078.269  2832.702     2758.636 855.007       4925.205
```{r}

cat("mean loss\n")
cbind(MICEloss = CCMICEmeanloss, MRloss = CCMRmeanloss, MIDASloss = CCMIDASmeanloss, MIDASVAEloss = CCMIDASVAEmeanloss, CCLoss=CCmeanloss)

cat("\nsum loss\n")
cbind(MICEloss = CCMICEsumloss, MRloss = CCMRsumloss, MIDASloss = CCMIDASsumloss, MIDASVAEloss = CCMIDASVAEsumloss, CCLoss=CCsumloss)

```

```{r}
CCMICEbrier = mean((CCMICEpreds - outcome)^2)
CCMRbrier = mean((CCMRpreds - outcome)^2)
CCMIDASbrier = mean((CCMIDASpreds - outcome)^2)
CCMIDASVAEbrier = mean((CCMIDASVAEpreds - outcome)^2)
CCbrier = mean((CCpreds - CCoutcome)^2)


cat("\nBrier scores\n")
cbind(CCMICEbrier, CCMRbrier, CCMIDASbrier, CCMIDASVAEbrier, CCbrier, adjustedCCbrier)

```

Compared to
Brier scores
      MICEbrier    MRbrier MIDASbrier MIDASVAEbrier    CCbrier adjustedCCbrier
[1,] 0.06899898 0.08221635 0.07204966    0.06947581 0.05159532       0.1409384


```{r}
showmetrics(CCMICEpreds, outcome)
showmetrics(CCMRpreds, outcome)
showmetrics(CCMIDASpreds, outcome)
showmetrics(CCMIDASVAEpreds, outcome)
showmetrics(adjustedCCpreds, outcome)
showmetrics(CCpreds, CCoutcome)

```

```{r}
CCMICEroc = roc(outcome~CCMICEpreds)
(auc(CCMICEroc))
CCMICEci = ci(CCMICEroc)

CCMIDASroc = roc(outcome~CCMIDASpreds)
(auc(CCMIDASroc))
CCMIDASci = ci(CCMIDASroc)


cmmice = confusionMatrix(as.factor( round(MICEpreds)),as.factor(outcome), positive='1')
cmmr = confusionMatrix(as.factor( round(MRpreds)),as.factor(outcome), positive='1')
cmmidas = confusionMatrix(as.factor( round(MIDASpreds)),as.factor(outcome), positive='1')
cmmidavaes = confusionMatrix(as.factor( round(MIDASVAEpreds)),as.factor(outcome), positive='1')
cmcc = confusionMatrix(as.factor( round(CCpreds)),as.factor(CCoutcome), positive='1')
#use this as base
cmccmice = confusionMatrix(as.factor( round(CCMICEpreds)),as.factor(outcome), positive='1')
cmccmidas = confusionMatrix(as.factor( round(CCMIDASpreds)),as.factor(outcome), positive='1')


# cmmice$byClass["Pos Pred Value"]
#   overall["Accuracy"]


method = c("Complete case (MIDAS Test Data)", "MICE", "missRanger", "MIDAS", "MIDASVAE")
metric = c("Accuracy", "Brier", "Pos Pred Value", "Neg Pred Value", "Mean Loss", "AUC")
accuracy = c(cmccmidas$overall["Accuracy"],cmmice$overall["Accuracy"],cmmr$overall["Accuracy"]
             ,cmmidas$overall["Accuracy"], cmmidasvae$overall["Accuracy"])
accuracy = accuracy/accuracy[1] * 100

Brier = c(CCMICEbrier, MICEbrier, MRbrier, MIDASbrier, MIDASVAEbrier)
Brier = Brier/Brier[1] * 100

posPredValue = c(cmccmidas$byClass["Pos Pred Value"],cmmice$byClass["Pos Pred Value"],cmmr$byClass["Pos Pred Value"],cmmidas$byClass["Pos Pred Value"],cmmidasvae$byClass["Pos Pred Value"])
posPredValue =   posPredValue/posPredValue[1] * 100

negPredValue = c(cmccmidas$byClass["Neg Pred Value"],cmmice$byClass["Neg Pred Value"],cmmr$byClass["Neg Pred Value"],cmmidas$byClass["Neg Pred Value"],cmmidasvae$byClass["Neg Pred Value"])
negPredValue = negPredValue/negPredValue[1] * 100

meanloss = c(CCMIDASmeanloss, MICEmeanloss, MRmeanloss, MIDASmeanloss, MIDASVAEmeanloss)
meanloss = meanloss/meanloss[1] * 100

auc = c(CCMIDASci[2], MICEci[2],MRci[2], MIDASci[2], MIDASVAEci[2])
auc = auc/auc[1] * 100
# CCci[1]
# CCci[2]
# CCci[3]
# CCci
#summarydf = data.frame(method, accuracy, accuracyN, precision, precisionN, negpredvalue, negpredvalueN, meanloss, meanlossN, auc,
summarydf = data.frame(method, accuracy, Brier, posPredValue, negPredValue, meanloss, auc)
```

```{r}
library(reshape)
library(data.table)

sdf2 = data.frame(t(summarydf))
sdf2 

names(sdf2) = summarydf$method
setDT(sdf2, keep.rownames = TRUE)[]
colnames(sdf2)[1] <- "Metric"

#sdf3 = sdf2[c(3,5,7,9, 11),]
sdf3 = sdf2[-1]
sdf3$"Complete case (MIDAS Test Data)"=round(as.numeric(as.character(sdf3$"Complete case (MIDAS Test Data)")),1)
sdf3$"MICE"=round(as.numeric(as.character(sdf3$"MICE")),1)
sdf3$"missRanger"=round(as.numeric(as.character(sdf3$"missRanger")),1)
sdf3$"MIDAS"=round(as.numeric(as.character(sdf3$"MIDAS")),1)
sdf3$"MIDASVAE"=round(as.numeric(as.character(sdf3$"MIDASVAE")),1)
sdf3

```

```{r}
library(reshape2)

# reshape your data into long format
sdf3l <- melt(sdf3, id=c("Metric"))
sdf3l

#create color palette:
library(RColorBrewer)
coul = brewer.pal(5, "Dark2") 

#values = c("red","blue","green","yellow"), 
# make the plot
ggplot(sdf3l) +
  geom_bar(aes(x = Metric, y = value, fill = variable), 
           stat="identity", position = "dodge", width = 0.7) +
  scale_fill_manual("Method\n", values = coul, 
                    labels = c("Complete Case\n(MIDAS Test Data)", "MICE", "missRanger", "MIDAS", "MIDASVAE")) +
  labs(x="",y="Relative Performance %\n") +
  ggtitle("Performance Relative to Complete Case with MIDAS (100%)") +
  ylim(c(0,160)) +
  theme_bw(base_size = 14) + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

