---
title: "R Notebook"
output: html_notebook
---
This version with some ASA transformations incorporated
```{r}
parentfolder = "c:\\PDH\\Moh\\"
datafile = "MOHFinalV3.csv"

df<- read.csv(paste0(parentfolder,datafile), header=TRUE
              ,comment.char = "",  skip =0,  check.names = FALSE
              ,na.strings=c("NA","NaN", ""))#,colClasses=mfColInfo)
str(df)
summary(df)
```


```{r}
# Minor Cleanup and Transformations
df$ethnicGroup=as.factor(df$ethnicGroup)
df$DHB=as.factor(df$DHB)
#df$facility=as.factor(df$facility)
df$opChapNum=as.factor(df$opChapNum)
df$op02ChapNum=as.factor(df$op02ChapNum)
df$op03ChapNum=as.factor(df$op03ChapNum)
df$opSeverity=as.factor(df$opSeverity)
df$ASA=as.factor(df$ASA)
df$isEmergency=as.factor(df$isEmergency)
df$InHospitalMortality=as.factor(df$InHospitalMortality)
df$ShortTermMortality1d=as.factor(df$ShortTermMortality1d)
df$ShortTermMortality3d=as.factor(df$ShortTermMortality3d)
df$LongTermMortality30d=as.factor(df$LongTermMortality30d)
df$LongTermMortality1year=as.factor(df$LongTermMortality1year)
df$LongTermMortality2year=as.factor(df$LongTermMortality2year)
df$Dep_score <- as.ordered(df$Dep_score)
df$ASA=as.ordered(df$ASA)

# Some non ascii characters are present in the 1st column. Rename to clear up
colnames(df)[1] <- "gender"

# ASA 9 means never got one. For our purposes, we treat this the same as missing.
 df[which((df$ASA)==9),]$ASA=NA 
#df.data = df[,c(-10, -22, -25, -28, -31, -36)]
#Omit off drugs1m, labs1m etc. We only want 6m and 12m. Take off mortality1d as too much overlap.
df.data = df[,c(-22, -25, -28, -31, -36)] 
str(df.data)

#apply(df.data,2, class)
summary(df.data)
#summary (df.data$ASA)
#summary (df.data$opChapNum)

datafile = "MOHFinalV3Cleaned.csv"
write.csv(df.data, paste0(parentfolder,datafile), row.names=F,  na = "")

```
Additional processing to simplify the dataset
and also to align it with requirements for NZ Surgical Risk calculator.
The NZ Surgical Risk calculator uses the following fields

1) Acuity             - Acute or not acute
2) Age                - Continuous integer
3) Cancer Status      - A true / false
4) ASA                - Using ASA as a 3 group category. ASA1/2 , ASA 3 , ASA 4/5
5) Grade of surgery   - Factor of 1-5
6) Surgical specialty - High risk named specialties and other. 
The last two are suggested as 'additional predictors' in the plan
7) Gender             - Male / Female
8) Ethnicity          - Using 3 levels, Maori / Pacific / Other 


Preprocessing to make it consistent with previous work
```{r}
#not required
#d_covariates = d_covariates[-(which(d_covariates$opSeverity == 999)),]
#d_covariates$opSeverity=d_covariates$opSeverity[,drop=TRUE]
d_covariates = df.data
#summary(d_covariates)
#drop ASA 6 since they are already dead before hitting the table
d_covariates = d_covariates[-(which(d_covariates$ASA == 6)),]
d_covariates$ASA=d_covariates$ASA[,drop=TRUE]
#d_covariates$ASA = as.factor(d_covariates$ASA)
#d_covariates$ASA = as.ordered(d_covariates$ASA)

#Leave this for later
# summary(d_covariates$ASA)
# d_covariates$threeGroupASA = ifelse(d_covariates$ASA == 1, yes = "ASA1/2"
#                                 , no = ifelse(d_covariates$ASA == 2, yes = "ASA1/2"
#                                     , no = ifelse(d_covariates$ASA == 3, yes = "ASA 3"
#                                         , no = ifelse(d_covariates$ASA == 4, yes = "ASA 4/5"
#                                             , no = "ASA 4/5"))))
# 
# d_covariates$threeGroupASA = relevel(as.factor(d_covariates$threeGroupASA), ref = "ASA1/2")

## Things that need to be classified from the database dataset. 
# These were all tested in multiple ways and then decided on as the final pass way of including them

## Ethnicity is used as a 3 level covariate. Maori, Pacific, Not Maori/Pacific. 
d_covariates$ethnicity3Groups = NA
#nrow(d_covariates)
d_covariates$ethnicity3Groups[which(d_covariates$ethnicGroup == 2 | d_covariates$ethnicGroup == 21)] = 'Maori'
d_covariates$ethnicity3Groups[which(d_covariates$ethnicGroup == 3 | (as.numeric(as.character(d_covariates$ethnicGroup))>= 30 & as.numeric(as.character(d_covariates$ethnicGroup)) <= 37))] = 'Pacific'
d_covariates$ethnicity3Groups[which(is.na(d_covariates$ethnicity3Groups))] = 'Not Maori/Pacific'
# Make sure to relevel factors so that the odds ratios appear in the right order
d_covariates$ethnicity3Groups = relevel(as.factor(d_covariates$ethnicity3Groups), ref = 'Not Maori/Pacific')

# Grade needs to be converted into a two factor variable moving 4/5 together vs the others
# d_covariates$Surg_grade_4_5 = ifelse(d_covariates$Surgical_grade == 5, yes = "4_5", no = ifelse(d_covariates$Surgical_grade == 4, yes = "4_5", no = "1-3"))
d_covariates$Surg_grade_4_5 = ifelse(d_covariates$opSeverity == 5, yes = "4_5"
                                     , no = ifelse(d_covariates$opSeverity == 4, yes = "4_5"
                                     , no = ifelse(is.na(d_covariates$opSeverity) == TRUE, yes = NA, no= "1-3")))

#check this
d_covariates$Surg_grade_4_5 = relevel(as.factor(d_covariates$Surg_grade_4_5), ref = "1-3")

summary(d_covariates$Surg_grade_4_5)
#is.na(d_covariates$opSeverity) 
d_covariates$Acuity = ifelse(d_covariates$AdmissionType == "AC", yes = "acute", no = "elective")
d_covariates$Acuity = relevel(as.factor(d_covariates$Acuity), ref = "elective")

d_covariates$Age = as.integer(floor(d_covariates$opAgeYearsFractional))

d_covariates$cancerStatus = as.logical(ifelse(d_covariates$IsCancer == "Y", yes = "TRUE", no = "FALSE"))



# # Need to set up the death times from the data - used for different outlook models (30d, 1y, 2y)
# d_covariates$deathAt30 = ifelse(timeToDeath < 31, yes = 1, no = 0)
# d_covariates$deathAt365 = ifelse(timeToDeath < 366, yes = 1, no = 0)
# d_covariates$deathAt730 = ifelse(timeToDeath < 731, yes = 1, no = 0)
d_covariates$Mortality30 = d_covariates$LongTermMortality30d
d_covariates$Mortality365 = d_covariates$LongTermMortality1year
d_covariates$Mortality730 = d_covariates$LongTermMortality2year#


# Giving surgical specialties a name rather than a number for cleaner output
# Chapter 1 Procedures on Nervous System [1-86]
# Chapter 2 Procedures on Endocrine System [110-129]
# Chapter 3 Procedures on Eye and Adnexa [160-256]
# Chapter 4 Procedures on Ear and Mastoid Process [300-333]
# Chapter 5 Procedures on Nose, Mouth and Pharynx [370-422]
# Chapter 6 Dental Services [450-490]
# Chapter 7 Procedures on Respiratory System [520-571]
# Chapter 8 Procedures on Cardiovascular System [600-777]
# Chapter 9 Procedures Blood and Blood-Forming Organs [800-817]
# Chapter 10 Procedures on Digestive System [850-1011]
# Chapter 11 Procedures on Urinary System [1040-1129]
# Chapter 12 Procedures on Male Genital Organs [1160-1203]
# Chapter 13 Gynaecological Procedures [1240-1299]
# Chapter 14 Obstetric Procedures [1330-1347]
# Chapter 15 Procedures on Musculoskeletal System [1360-1580]
# Chapter 16 Dermatological and Plastic Procedures [1600-1718]
# Chapter 17 Procedures on Breast [1740-1759]
# Chapter 18 Radiation Oncology Procedures [1786-1799]
# Chapter 19 Noninvasive, cognitive and Other Interventions, Not Elsewhere Classified [1820-1922]
# Chapter 20 Imaging Services [1940-2016] 
summary(d_covariates$opChapNum)
summary(d_covariates$op02ChapNum)
summary(d_covariates$op03ChapNum)

#Do this post imputation
# fullSurgSpecList = c("1_nervousSystem","2_endocrineSystem","3_eyeAndAdnexa","4_earAndMastoidProcess","5_noseMouthAndPharynx","7_respiratorySystem","8_vascularSystem"
#                      ,"9_bloodAndBloodFormingOrgans","10_digestiveSystem","11_urinarySystem","12_maleGenitalOrgans","13_gynaecological","14_obstetric","15_musculoskeletalSystem"
#                      ,"16_dermatologicalAndPlastic")
# fullSurgSpecList2 = c("0_None","1_nervousSystem","2_endocrineSystem","3_eyeAndAdnexa","4_earAndMastoidProcess","5_noseMouthAndPharynx","6_dental","7_respiratorySystem"
#                       ,"8_vascularSystem","9_bloodAndBloodFormingOrgans","10_digestiveSystem","11_urinarySystem","12_maleGenitalOrgans","13_gynaecological","14_obstetric"
#                       ,"15_musculoskeletalSystem","16_dermatologicalAndPlastic","18_radiationOncology","19_NonInvasiveandOther","20_Imaging")
# fullSurgSpecList3 = c("0_None","1_nervousSystem","2_endocrineSystem","3_eyeAndAdnexa","4_earAndMastoidProcess","5_noseMouthAndPharynx","6_dental","7_respiratorySystem"
#                       ,"8_vascularSystem","9_bloodAndBloodFormingOrgans","10_digestiveSystem","11_urinarySystem","12_maleGenitalOrgans","13_gynaecological","14_obstetric"
#                       ,"15_musculoskeletalSystem","16_dermatologicalAndPlastic","17_breast","18_radiationOncology","19_NonInvasiveandOther","20_Imaging")

#d_covariates$surgSpecBreakdownFullList = mapvalues(x=d_covariates$Surg_spec, from = levels(as.factor(d_covariates$Surg_spec)), to=fullSurgSpecList)
# d_covariates$surgSpec01 = mapvalues(x=d_covariates$opChapNum, from = levels(as.factor(d_covariates$opChapNum)), to=fullSurgSpecList)
# d_covariates$surgSpec02 = mapvalues(x=d_covariates$op02ChapNum, from = levels(as.factor(d_covariates$op02ChapNum)), to=fullSurgSpecList2)
# d_covariates$surgSpec03 = mapvalues(x=d_covariates$op03ChapNum, from = levels(as.factor(d_covariates$op03ChapNum)), to=fullSurgSpecList3)

summary(d_covariates)
#remove enthnicGrop, DHB, endType, opSeverity, opAgeYears,opAgeYearsFractional, AdmissionType, AgeAtAdmission, ShortermMortality3d, 30d, 1year, 2year
df.data2 = d_covariates[,c(-2,-3,-5,-9, -10, -11, -13, -16, -30, -32, -33, -34, -35)]
summary(df.data2)

cat(toString(shQuote(sapply(df.data2, class))), "\n")
str(df.data2)
datafile = "MOHFinalV4Cleaned.csv"
write.csv(df.data2, paste0(parentfolder,datafile), row.names=F,  na = "")
```

#################################################################################################
```{r}
#Original - if we plan to split before imputing
dfOrg = df.data
set.seed(101)
sample <- sample.int(n = nrow(dfOrg), size = floor(.70*nrow(dfOrg)), replace = F)
trainOrg <- dfOrg[sample, ]
testOrg  <- dfOrg[-sample, ]

par(mfrow=c(1,2))
plot(trainOrg$InHospitalMortality)
plot(testOrg$InHospitalMortality)

write.csv(trainOrg, paste0(parentfolder,"MOHFinalV4OrgTrain.csv"), row.names=F,  na = "") 
write.csv(testOrg, paste0(parentfolder,"MOHFinalV4OrgTest.csv"), row.names=F,  na = "")  
```


Create Complete Case Train and Test Dataset where records with missing values are dropped.

```{r}
dfComplete = na.omit(df.data)
summary(dfComplete)

set.seed(101)
sample <- sample.int(n = nrow(dfComplete), size = floor(.70*nrow(dfComplete)), replace = F)
trainNM <- dfComplete[sample, ]
testNM  <- dfComplete[-sample, ]

par(mfrow=c(1,2))
plot(trainNM$InHospitalMortality)
plot(testNM$InHospitalMortality)

write.csv(trainNM, paste0(parentfolder,"MOHFinalV4CompleteCaseTrain.csv"), row.names=F,  na = "")  #11766
write.csv(testNM, paste0(parentfolder,"MOHFinalV4CompleteCaseTest.csv"), row.names=F,  na = "")  #5043

# dfNoMissingNo9s = dfNoMissing[which((dfNoMissing$ASA)!=9),]
# set.seed(101)
# sample <- sample.int(n = nrow(dfNoMissingNo9s), size = floor(.70*nrow(dfNoMissingNo9s)), replace = F)
# trainNMN9 <- dfNoMissingNo9s[sample, ]
# testNMN9  <- dfNoMissingNo9s[-sample, ]
# write.csv(trainNMN9, "c:\\PDH\\MoH18PlusTrainMissing9Dropped.csv", row.names=F,  na = "")  #10002
# write.csv(testNMN9, "c:\\PDH\\MoH18PlusTestMissing9Dropped.csv", row.names=F,  na = "")    #4287
```


Play around with correlation
```{r}
# dfComplete$InHospitalMortality=as.integer(dfComplete$InHospitalMortality)
# nums <- unlist(lapply(dfComplete, is.numeric))  
# dta.r <- abs(cor(dfComplete[ , nums]))
# dta.r
# dfComplete$InHospitalMortality=as.factor(dfComplete$InHospitalMortality)
```

Manual Impute. Modes and Means.
Create Manual Impute Train and Test Datasets.
```{r}
dfImputed = df.data

Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

dfImputed[which(is.na(dfImputed$opSeverity)),]$opSeverity=2 #2 what about 999? Filtered by noncardiac events

#dfImputed[which((dfImputed$ASA)==9),]$ASA=2 #11895
#dfImputed[which(is.na(dfImputed$ASA)),]$ASA=Mode(dfImputed$ASA) #7739
dfImputed[which(is.na(dfImputed$ASA)),]$ASA = 2

dfImputed[which(is.na(dfImputed$isEmergency)),]$isEmergency=Mode(dfImputed$isEmergency)

dfImputed$Dep_score <- as.ordered(dfImputed$Dep_score)
dfImputed[which(is.na(dfImputed$Dep_score)),]$Dep_score=Mode(dfImputed$Dep_score)

summary(dfImputed)

set.seed(101)
sample <- sample.int(n = nrow(dfImputed), size = floor(.70*nrow(dfImputed)), replace = F)
trainImpute <- dfImputed[sample, ]
testImpute  <- dfImputed[-sample, ]
par(mfrow=c(1,2))
plot(trainImpute$InHospitalMortality)
plot(testImpute$InHospitalMortality)

write.csv(trainImpute, paste0(parentfolder,"MOHFinalV4ImputeTrain.csv"), row.names=F,  na = "")  #27584
write.csv(testImpute, paste0(parentfolder,"MOHFinalV4ImputeTest.csv"), row.names=F,  na = "")    #11823

```

If we had to reload.
```{r}
# mfColInfo2 = c("factor", "factor", "factor", "factor", "factor", "factor", "factor", "factor", "factor", "numeric"
#                ,"integer", "factor", "factor", "factor", "factor", "factor", "factor", "factor", "factor", "factor"
#                ,"integer", "integer", "integer", "integer", "integer", "integer", "integer", "integer", "numeric"
#                ,"factor","factor", "factor", "factor", "factor")
# datafile2 = paste0(parentfolder, "FirstAdmissions18PlusFrom2006SubsetDistilledV2NCCleaned.csv")
# 
# df2<- read.csv(datafile2, header=TRUE
#               ,comment.char = "",  skip =0,  check.names = FALSE
#               ,na.strings=c("NA","NaN", ""),colClasses=mfColInfo2)
# df2$Dep_score <- as.ordered(df2$Dep_score)
# 
# #df2[which((df2$ASA)==9),]$ASA=NA #11895
# df.dataForMI = df2
# summary(df.dataForMI)
```

Do the missForest MI imputation

```{r}
df.dataForMI = df.data
#install.packages("doParallel")
#install.packages("missForest")
require(doParallel)
registerDoParallel(cores=4)
getDoParWorkers()
library(missForest)

set.seed(81)
mffullfd.start = Sys.time()
imp.mffullforests = missForest(df.dataForMI, variablewise = TRUE, verbose = TRUE, parallelize = 'forests')
mffullfd.end = Sys.time()
mffull.forestsduration = mffullfd.end - mffullfd.start
mffull.forestsduration

#Time difference of 1.530393 days Mimic
# Time difference of 11.49234 mins - Time difference of 4.994217 mins

summary(imp.mffullforests$ximp)
str(imp.mffullforests$ximp)
summary(imp.mffullforests$OOBerror)
imp.mffullforests$OOBerror

saveRDS(imp.mffullforests, file = paste0(parentfolder,"MOHFinalV4MFMI.rds") )
write.csv(imp.mffullforests$ximp, file = paste0(parentfolder,"MOHFinalV4MFMI.csv"), row.names=F)
write.csv(imp.mffullforests$OOBerror, file = paste0(parentfolder,"MOHFinalV4MFMIOOBerror.csv") )

save.image(file = paste0(parentfolder, "moh_work_space_9020.RData"))

```


IF we had to reload else comment this out
```{r}
# mfColInfo3 = c("factor", "factor", "factor", "factor", "factor", "factor", "factor", "factor", "factor", "numeric"
#                ,"integer", "factor", "factor", "factor", "factor", "factor", "factor", "factor", "factor", "factor"
#                ,"integer", "integer", "integer", "integer", "integer", "integer", "integer", "integer", "numeric"                 ,"factor","factor", "factor", "factor", "factor")
# 
# imputeddatafile = paste0(parentfolder, "mohimp.mffullforests.csv")
# 
# dfmi<- read.csv(imputeddatafile, header=TRUE
#                 ,comment.char = "",  skip =0,  check.names = FALSE
#                 ,na.strings=c("NA","NaN", ""),colClasses=mfColInfo3)
# 
# 
# summary(dfmi)
# str(dfmi)
```


Now create the MI test and train sets.
```{r}
dfmi = imp.mffullforests$ximp
set.seed(101)
sample <- sample.int(n = nrow(dfmi), size = floor(.70*nrow(dfmi)), replace = F)
trainMI <- dfmi[sample, ]
testMI  <- dfmi[-sample, ]
par(mfrow=c(1,2))
plot(trainMI$InHospitalMortality)
plot(testMI$InHospitalMortality)

write.csv(trainMI, paste0(parentfolder,"MOHFinalV4MITrain.csv"), row.names=F,  na = "")  #
write.csv(testMI, paste0(parentfolder,"MOHFinalV4MITest.csv"), row.names=F,  na = "")    #

```
#if we make the ASA 3 level before imputing...
```{r}
require(doParallel)
registerDoParallel(cores=4)
getDoParWorkers()
library(missForest)


df.dataForMI2 = df.data
# To be consistent, drop ASA 6 since they are already dead before hitting the table
df.dataForMI2 = df.dataForMI2[-(which(df.dataForMI2$ASA == 6)),]
df.dataForMI2$ASA=df.dataForMI2$ASA[,drop=TRUE]

df.dataForMI2$ASA = ifelse(df.dataForMI2$ASA == 1, yes = "ASA1/2"
                                , no = ifelse(df.dataForMI2$ASA == 2, yes = "ASA1/2"
                                    , no = ifelse(df.dataForMI2$ASA == 3, yes = "ASA 3"
                                        , no = ifelse(df.dataForMI2$ASA == 4, yes = "ASA 4/5"
                                            , no = "ASA 4/5"))))
#df.dataForMI2$ASA = as.factor(df.dataForMI2$ASA)
df.dataForMI2$ASA = relevel(as.factor(df.dataForMI2$ASA), ref = "ASA1/2")

summary(df.dataForMI2)
set.seed(81)
mffullfd2.start = Sys.time()
imp.mffullforests2 = missForest(df.dataForMI2, variablewise = TRUE, verbose = TRUE, parallelize = 'forests')
mffullfd2.end = Sys.time()
mffull.forestsduration2 = mffullfd2.end - mffullfd2.start
mffull.forestsduration2

#Time difference of 1.530393 days Mimic vs Time difference of 5.900927 mins

summary(imp.mffullforests2$ximp)
str(imp.mffullforests2$ximp)
summary(imp.mffullforests2$OOBerror)
imp.mffullforests2$OOBerror

saveRDS(imp.mffullforests2, file = paste0(parentfolder,"MOHFinalV4MFMI3LvlASA.rds") )
write.csv(imp.mffullforests2$ximp, file = paste0(parentfolder,"MOHFinalV4MFMI3LvlASA.csv"), row.names=F)
write.csv(imp.mffullforests2$OOBerror, file = paste0(parentfolder,"MOHFinalV4MFMI3LvlASAOOBerror.csv") )

save.image(file = paste0(parentfolder, "moh_work_space_3LvlASA.RData"))
 #     ASA        isEmergency  IsCancer 
 # 1   : 5193   Min.   : 18.00   Min.   : 18.00       Min.   :   0.000   AA: 7290      ASA1/2 :12908   0   : 5789   N:35958  
 # 2   :21931   1st Qu.: 36.00   1st Qu.: 35.94       1st Qu.:   0.000   AC:18469      ASA 3  : 2973   1   :25142   Y: 3449  
 # 3   : 8750   Median : 55.00   Median : 55.35       Median :   1.000   AP:    1      ASA 4/5: 1882   NA's: 8476            
 # 4   : 2880   Mean   : 54.54   Mean   : 54.53       Mean   :   3.361   RL:    1      NA's   :21644 


```

```{r}

#install.packages("mice")

require(mice)
require(lattice)

set.seed(81)
starttm = Sys.time()
mohimp <- parlmice(df.data2, m=40, maxit=40, cluster.seed = 81, print=TRUE)
endtm = Sys.time()
micetime = endtm - starttm
micetime
#Time difference of 2.197196 hours - actually moh
saveRDS(mohimp, file = paste0(parentfolder,"MiceImpMohV4Final.rds"))
```


