
```{r}
parentfolder = "c:\\PDH\\Moh\\"
datafile = "MOHFinalSecondClean.csv"

mColInfo = c("factor", "factor", "factor", "factor", "factor", "factor", "integer", "factor", "factor", "factor"
             , "factor", "factor", "factor", "factor", "integer", "integer", "integer", "integer", "integer", "integer"
             , "integer", "integer", "factor", "factor", "factor", "character", "integer", "logical", "factor", 
             "factor", "factor"  )

df<- read.csv(paste0(parentfolder,datafile), header=TRUE
              ,comment.char = "",  skip =0,  check.names = FALSE
              ,na.strings=c("NA","NaN", ""),colClasses=mColInfo)

df$ethnicity3Groups = relevel(as.factor(df$ethnicity3Groups), ref = 'Not Maori/Pacific')
#df$Surg_grade_4_5 = relevel(as.factor(df$Surg_grade_4_5), ref = "1-3")
df$Acuity = relevel(as.factor(df$Acuity), ref = "elective")
df$Dep_score=as.ordered(df$Dep_score)
df$ASA=as.ordered(df$ASA)

str(df)
summary(df)

# to avoid 1<10<2<3......
str(df$Dep_score)
df$Dep_score = as.numeric(as.character(df$Dep_score))
df$Dep_score = as.ordered(df$Dep_score)

# to avoid 1<10<2<3......
df$opChapNum = as.numeric(as.character(df$opChapNum))
df$opChapNum = as.factor(df$opChapNum)
df$op02ChapNum = as.numeric(as.character(df$op02ChapNum))
df$op02ChapNum = as.factor(df$op02ChapNum)
df$op03ChapNum = as.numeric(as.character(df$op03ChapNum))
df$op03ChapNum = as.factor(df$op03ChapNum)


#true binary
# isEmergency,InHospitalMortality,cancerStatus,Mortality30,Mortality365,Mortality730
# binary cats -> recode to binary
# ,gender_F,gender_M
# ,IsSmoker_N,IsSmoker_Y,WasSmoker_N,WasSmoker_Y,HasDiabetesT1_N,HasDiabetesT1_Y,HasDiabetesT2_N,HasDiabetesT2_Y,IsNeuroTrauma_N,IsNeuroTrauma_Y
# ordinal
# opSeverity,ASA,Dep_score
# True cats
# ,eventType_ID,eventType_IM,eventType_IP
# ,opChapNum,op02ChapNum,op03ChapNum,
# ,ethnicity3Groups_Maori,ethnicity3Groups_Not Maori/Pacific,ethnicity3Groups_Pacific
levels(df$IsSmoker) <- c(0,1)
levels(df$WasSmoker) <- c(0,1)
levels(df$HasDiabetesT1) <- c(0,1)
levels(df$HasDiabetesT2) <- c(0,1)
levels(df$IsNeuroTrauma) <- c(0,1)
levels(df$gender) <- c(0,1)
names(df)[1] <- "IsMale"
names(df)[names(df) == 'gender'] <- 'IsMale'
names(df)[names(df) == 'isEmergency'] <- 'IsEmergency'
names(df)[names(df) == 'cancerStatus'] <- 'HasCancer'
df$HasCancer = as.factor(df$HasCancer)
levels(df$HasCancer) <- c(0,1)




str(df)
summary(df)
table(df$Dep_score)

#Split into train and test datasets
#Original - if we plan to split before imputing
dfOrg = df
set.seed(101)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
```{r}
library(ggplot2)








```

```{r}
install.packages("naniar")
library(naniar)
vis_miss(airquality)
vis_miss(df)
```

```{r}
set.seed(101)
sample <- sample.int(n = nrow(dfOrg), size = floor(.70*nrow(dfOrg)), replace = F)
trainOrg <- dfOrg[sample, ]
testOrg  <- dfOrg[-sample, ]
```

```{r}
vis_miss(trainOrg)
```

```{r}
vis_miss(testOrg)
```
```{r}
gg_miss_upset(df)
```

```{r}
n_var_miss(df)
```

```{r}
gg_miss_upset(df, nsets = n_var_miss(df))
```
```{r}
gg_miss_upset(df, 
              nsets = 10,
              nintersects = NA)
```
```{r}
library(ggplot2)
# using regular geom_point()
ggplot(airquality,
       aes(x = Ozone,
           y = Solar.R)) +
geom_point()
```

```{r}
library(naniar)

# using  geom_miss_point()
ggplot(airquality,
       aes(x = Ozone,
           y = Solar.R)) +
 geom_miss_point()
```

```{r}
# Facets!
ggplot(airquality,
       aes(x = Ozone,
           y = Solar.R)) +
 geom_miss_point() + 
 facet_wrap(~Month)
```

```{r}
ggplot(df,
       aes(x = as.numeric(as.character(Dep_score)),
           y = LOS)) +
 geom_miss_point() + 
 facet_wrap(~IsSmoker)


ggplot(df,
       aes(x = as.numeric(as.character(Dep_score)),
           y = LOS)) +
 geom_miss_point() + 
 facet_wrap(~ASA)


ggplot(df,
       aes(x = as.numeric(as.character(opSeverity)),
           y = LOS)) +
 geom_miss_point() + 
 facet_wrap(~ASA)


ggplot(df,
       aes(x = as.numeric(as.character(ASA)),
           y = LOS)) +
 geom_miss_point() + 
 facet_wrap(~opSeverity)

```

```{r}

gg_miss_var(df)

library(ggplot2)
gg_miss_var(df) + labs(y = "Look at all the missing ones")

gg_miss_var(df, show_pct = TRUE)

```


```{r}
gg_miss_var(df,
            facet = Acuity, show_pct = TRUE)
```
```{r}
gg_miss_case(df)
```
```{r}
gg_miss_fct(x = df$opSeverity, fct = df$ASA)
```

